---
title: 'Homework 6: Linear Models'
author: "Natalie Boychuk"
date: "12/7/2020"
output: html_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library(rvest)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(patchwork)
set.seed(1)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1 

```{r}

  homicide_df = 
    read_csv("data/homicide-data.csv") %>% 
    mutate(
      city_state = str_c(city, state, sep = "_"),
      resolved = case_when(
        disposition == "Closed without arrest" ~ 0,
        disposition == "Open/No arrest" ~ 0,
        disposition == "Closed by arrest" ~ 1
        )) %>% 
    filter((city_state != "Tulsa_AL"),
           (city_state != "Phoenix_AZ"),
           (city_state != "Kansas City_MO"),
           (city_state != "Dallas_TX"),
           (victim_race %in% c("White", "Black"))) %>% 
    select(city_state, resolved, victim_age, victim_race, victim_sex) %>% 
    mutate(victim_age = as.numeric(victim_age)) 
  
```

```{r}
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "Baltimore_MD")

glm(resolved ~ victim_age + victim_race + victim_sex,
    data = baltimore_df, 
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>%  
  knitr::kable(digits = 3)

```

Across cities 
```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
mutate(
  models = 
    map(.x = data, ~glm(glm(resolved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial()))),
  results = map(models, broom::tidy) 
  ) %>% 
select(city_state, results) %>% 
 unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))

```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Comments on the plot: 


## Problem 2 

```{r}
birthweight_df = 
  read_csv("data/birthweight.csv") %>% 
  mutate(mrace = as.factor(mrace),
         malform = as.factor(malform)) 

## Look for missing data 

tail(birthweight_df,100)

head(birthweight_df,100)

skimr::skim(birthweight_df)

## it all looks okay to me. Variable names are already clean and columns look good 

```

Given background knowledge/research (I work for a lab that studies maternal/infant mortality and morbidity and I did a quick literature search [ex: (https://www.marchofdimes.org/complications/low-birthweight.aspx)]), I first hypothesize that premature is the primary cause of low birthweight. I further hypothesize that cigarette smoking, mother's weight at delivery, mother's age at delivery, maternal race, and family income might confound this relationship.

I would also like to be clear that I am not hypothesizing that race in and of itself is a cause of low birthweight; rather, structural inequalities associated with being a Black mother in the United States cause low birthweight in infants.


My first step is to test to see whether the distribution is approximately normal by plotting cigarette smoking on bwt: 
```{r}
birthweight_df %>% 
  ggplot(aes(x = gaweeks, y = bwt)) + geom_point()
```
Looks like a linear model will not work here, based on my above results. Good to know. I will use the "glm" function to reflect that the data follow a non-Gaussian distribution. 


```{r fitting the initial model }

glm(bwt ~ gaweeks + mrace + smoken + momage + fincome + smoken * ppwt, data = birthweight_df) %>%
  broom::tidy() %>% 
  knitr::kable()

```

From this initial model, it does not look like the mother's age at delivery or family income is adding anything to the model. Also, it looks like mother's pre-pregnancy weight is significant but not in interaction with smoking. For parsimony, I'll drop these predictors that were insignificant in the model using a backwards stepwise approach, removing one at a time. 

```{r}
glm(bwt ~ gaweeks + mrace + smoken + momage + fincome + ppwt, data = birthweight_df) %>%
  broom::tidy() %>% 
  knitr::kable()
```

Fincome is still not significant so I will also drop this variable. 

```{r}
glm(bwt ~ gaweeks + mrace + smoken + momage + ppwt, data = birthweight_df) %>%
  broom::tidy() %>% 
  knitr::kable()
```

I will also drop momage, which is not significant when accounting for the other variables. 

```{r}
bwmodel = glm(bwt ~ gaweeks + mrace + smoken + ppwt, data = birthweight_df)

bwmodel %>% 
  broom::tidy() %>% 
  knitr::kable()

```

Okay, I'm happy with this model! Now I'm going to plot the model residuals and fitted values. 

```{r plotting residuals against fitted values}
birthweight_df %>% 
  add_residuals(bwmodel) %>%
      ggplot(aes(x = resid)) + geom_density()

## Slightly skewed but not too bad. Mostly centered around 0. 

birthweight_df %>% 
  add_residuals(bwmodel) %>%
    add_predictions(bwmodel) %>% 
      ggplot(aes(x = pred, y = resid)) + geom_point()

# This looks mostly okay, since it's almost the same residual plot as what was shown in lecture. There does seem to maybe be something weird happening in the top-left of the plot but for the purposes of this course I'll move on.

```

Model 2 (from HW): 
```{r}
model2 = glm(bwt ~ blength + gaweeks, data = birthweight_df) 

```

Model 3 (from HW): 
```{r}
model3 = glm(bwt ~ bhead * blength * babysex, data = birthweight_df)
```


Now I'm going to compare my model with the 2 provided models using cross-validation. 

```{r warning:FALSE cross-validation for prob 2 }

cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
   mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    bwmodel = map(train, ~mgcv::gam(bwt ~ s(gaweeks + mrace + smoken + ppwt), data = birthweight_df)),
    model2 = map(train, ~mgcv::gam(bwt ~ s(blength + gaweeks), data = birthweight_df)),
    model3 = map(train, ~mgcv::gam(bwt ~ s(bhead * blength * babysex), data = birthweight_df))) %>%
  mutate(
    rmse_bwmodel = map2_dbl(bwmodel, test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(model3, test, ~rmse(model = .x, data = .y)))

predic_error = 
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

predic_error

## Clearly mine is performing the worst (sad!)

```


## Problem 3 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r fitting a simple model to the data}
model_weather = 
  lm(tmax ~ tmin, data = weather_df) 
```

```{r getting values for a single model first}

rsquare = 
model_weather %>% 
  broom::glance() %>% 
  select(r.squared)

logb0b1 = 
model_weather %>% 
  broom::tidy() %>%
  pivot_wider(term,
              names_from = "term",
              values_from = "estimate") %>%
    mutate(logvalue = 
           log(`(Intercept)` * tmin)) %>% 
              select(logvalue)


```

```{r drawing 5000 bootstrap samples and getting distributions}

bootstraps = 
  weather_df %>% 
  bootstrap(5000, id = "strap_number") %>% 
  mutate(
    fit = map(.x = strap, ~lm(tmax ~ tmin, data = .x))) %>% 
                mutate(rsquare = map(fit, broom::glance), 
                      logvalue = map(fit, broom::tidy)) %>% 
  unnest(rsquare, logvalue) %>% 
  select(strap_number, term, r.squared, estimate) %>% 
  mutate(term=str_replace(term,"\\(Intercept\\)","intercept"))

rsquarevalue_boostraps = 
  bootstraps %>% 
  select(strap_number, r.squared) %>% 
  ggplot(aes(x = r.squared)) + geom_density()

rsquarevalue_boostraps

logvalue_bootstraps = 
bootstraps %>% 
     pivot_wider(term,
              names_from = "term",
              values_from = "estimate") %>%
    unnest(intercept, tmin) %>% 
    mutate(logvalue = 
           log(intercept * tmin)) %>% 
  ggplot(aes(x = logvalue)) + geom_density()

logvalue_bootstraps

```

The rsquared value seems to be centered around 0.91 and is slightly left-skewed, but is otherwise fairly symmetrical. 

The logvalue variable has a more symmetrical distribution around 2.02. There may be a very slight skewness to the right but it's hard to tell.

```{r confidence intervals}
final_logvalue = 
bootstraps %>% 
     pivot_wider(term,
              names_from = "term",
              values_from = "estimate") %>%
    unnest(intercept, tmin) %>% 
    mutate(logvalue = 
           log(intercept * tmin)) %>% 
  summarize(
    ci_lower = quantile(logvalue, 0.025), 
    ci_upper = quantile(logvalue, 0.975)) %>% 
    knitr::kable()

final_logvalue

final_rsquared = 
  bootstraps %>% 
  select(strap_number, r.squared) %>% 
  summarize(
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975)) %>% 
  knitr::kable()

final_rsquared

```


